Hydra Battles and Cie

### 2020-11-25 doc/hydras.tex
modifications in doc and README

\documentclass[a4paper]{book}
\usepackage{alltt}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{threeparttable}
\setcounter{tocdepth}{1}
\input{coqmacros}

@@

\section{Acknowledgements}
\label{sec:orgheadline5}
    Many thanks to Yves Bertot, David Ilcinkas, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
 members of the \emph{Formal Methods} team at laBRI for their helpful comments 
on an oral presentation of this work.

@@

\section{Index}

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main types and type classes}
\begin{tabular}{|c|c|c|l|}
\hline
Name & Description& Page \\\hline
Battle & type class of battle types & \pageref{types:Battle} \\
E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
ON & ordinal  notations & \pageref{types:ON} \\
ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\

@@

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main functions and constants}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina & Math & Description & Page \\ \hline
canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\

@@

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Main predicates}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
\hline
\end{tabular}
\begin{tablenotes}
  \item[1] This order is total, but not well-founded, because of not well formed terms.
\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
\item[3] This order is total \emph{and} well-founded.
\end{tablenotes}
\end{threeparttable}
\end{table}

\vspace{4pt}
\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Infix notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\

@@

\begin{table}[h]
  \centering
  \begin{threeparttable}
    \caption{Other notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
\pageref{sect:notation-FS}\\ 
omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

@@

% \section{Main notations}

% \subsubsection*{Intervals}

% Let $\leq$ be a total order on a type $A$. The intervals $\{k\in A| i\leq k\leq j\}$ is written $[i,j]$. Likewise, $\{k\in A| i\leq k<j\}$ is written
% $[i,j[$.

>>>>

\documentclass[a4paper]{book}
\usepackage{alltt}
\usepackage{makeidx}
\usepackage{hyperref}
\usepackage{url}
\usepackage{mdframed}
\usepackage{amsmath}
\usepackage{varioref}
\usepackage{synttree, proof, centernot}
\usepackage[firstpage]{draftwatermark}
\usepackage{verbatim}
\usepackage[note]{marginote}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{float}
\usepackage{threeparttable}
\setcounter{tocdepth}{1}
\input{coqmacros}

@@

\section{Acknowledgements}
\label{sec:orgheadline5}
    Many thanks to Yves Bertot, David Ilcinkas, Karl Palmskog, Sylvain Salvati, Alan Schmitt and Théo Zimmerman for their help on the elaboration of this document, and to the
 members of the \emph{Formal Methods} team at laBRI for their helpful comments 
on an oral presentation of this work.

@@

\chapter*{Index and tables}

\begin{todo}
 Still very incomplete!
\end{todo}

\section{Hydras and Ordinals}

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Main types and type classes}
\begin{tabular}{|c|c|c|l|}
\hline
Name & Description& Page \\\hline
Battle & type class of battle types & \pageref{types:Battle} \\
E0 & (well formed) ordinals below $\epsilon_0$ & \pageref{types:E0}\\
Hvariant & variants for proving termination of hydra battles & \pageref{sect:hvariant-def} \\
Hydra & tree-like representation of hydras & \pageref{types:Hydra} \\
Hydrae & finite sequences of hydras & \pageref{types:Hydrae} \\
ON & ordinal  notations & \pageref{types:ON} \\
ON\_for & comparison of an ordinal notation with Schütte's model &  \pageref{types:ON-for}\\
ON\_iso & isomorphism of ordinal notations  & \pageref{types:ON-iso}\\

@@

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Main functions and constants}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina & Math & Description & Page \\ \hline
canon & canon alpha i & $\canonseq{\alpha}{i}$& Canonical sequence & \pageref{Functions:canon}\\
canonS & canonS alpha i & $\canonseq{\alpha}{i+1}$& Helper for canon & \pageref{Functions:canonS}\\
F\_ & F\_ alpha n & $F_\alpha(n)$ & Wainer's \texttt{F} fast growing hierarchy & \pageref{Functions:F-alpha} \\
H\_ & H\_ alpha n & $H_\alpha(n)$ & Hardy's \texttt{H} fast growing hierarchy & \pageref{Functions:H-alpha} \\
iterate & iterate $f$ $n$ $x$ & $f^{(n)}(x)$ & Functional iteration & \pageref{Functions:iterate} \\
L\_ & L\_ alpha k & $L_\alpha(k)$ & final step of a standard path & \pageref{Functions:L-alpha} \\
succ & succ alpha & & Successor & \pageref{Functions:succ-T1}, \pageref{Functions:succ-sch} \dots \\
zero:Ord & & $0$ & The least ordinal (Schütte's model) & \pageref{Constants:zero:Ord} \\

@@

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Main predicates}
\begin{tabular}{|r | c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
\texttt{lt : T1->T1->Prop}& lt alpha beta & $\alpha < \beta$& strict order on type \texttt{T1} \tnote{1} & \pageref{Predicates:lt-T1}\\
\texttt{LT: T1->T1->Prop}& alpha o< beta & $\alpha < \beta$& strict order on type \texttt{T1}   \tnote{2} & \pageref{Predicates:LT-T1}\\
\texttt{Lt : E0->E0->Prop} & alpha o< beta & $\alpha < \beta$& strict order on type \texttt{E0} \tnote{3} & \pageref{Predicates:Lt-E0} \\
\texttt{nf: T1->Prop} & \texttt{nf alpha} && alpha is in Cantor normal form & \pageref{Predicates:nf-T1}\\
\hline
\end{tabular}
\begin{tablenotes}
  \item[1] This order is total, but not well-founded, because of not well formed terms.
\item[2] Restriction of \texttt{lt} to terms in normal form; this order is partial, but well-founded.
\item[3] This order is total \emph{and} well-founded.
\end{tablenotes}
\end{threeparttable}
\end{table}

\vspace{4pt}
\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Infix notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
 \texttt{on\_lt} & \texttt{alpha o< beta}&$\alpha<\beta$& ordinal inequality \tnote{1} & \pageref{sect:on-lt-notation}\\
 \texttt{on\_le} & \texttt{alpha o<= beta}&$\alpha\leq\beta$& ordinal inequality & \pageref{sect:on-lt-notation}\\
\texttt{plus} & \texttt{alpha + beta} & $\alpha + \beta$ & ordinal addition & \pageref{sect:infix-plus-T1}, \dots\\
\texttt{oplus} & \texttt{alpha o+ beta} & $\alpha \oplus \beta$ & Hessenberg sum & \pageref{sect:infix-oplus} \\
\texttt{round} & \texttt{h -1-> h'} & & one round of a battle & \pageref{sect:infix-round} \\
\texttt{rounds} & \texttt{h -+-> h'} & & one or more  rounds of a battle & \pageref{sect:infix-rounds} \\
\texttt{round\_star} & \texttt{h -*-> h'} & & any number of rounds of a battle & \pageref{sect:infix-rounds} \\

@@

\begin{table}[H]
  \centering
  \begin{threeparttable}
    \caption{Other notations}
\begin{tabular}{|c|c|c|c|l|}
\hline
Name & Gallina&Math& Description& Page \\\hline
F & \texttt{F $n$} & $n$ & The $n$-th finite ordinal &  
\pageref{sect:notation-F}, \pageref{sect:notation-F-sch}\\ 
FS & \texttt{FS $n$} & $n+1$ & The $n+1$-th finite ordinal  \tnote{2} &  
\pageref{sect:notation-FS}\\ 
omega & \texttt{omega} & $\omega$ &   the first infinite ordinal   & \pageref{sect:notation-omega}, \pageref{sect:omega-T1}, \pageref{sect:omega-notation2}, \dots\\
phi0     & \texttt{phi0 alpha} & $\phi_0(\alpha),\; \omega^\alpha$&exponential of base $\omega$ & \pageref{sect:notation-phi0}\\

@@

\section{Addition Chains}

\begin{todo}
\end{todo}
